\documentclass[10pt]{report}

\usepackage{appendix}
\usepackage{floatflt}
\usepackage{fancyhdr}
\usepackage{textcomp}
\usepackage[usenames]{color}
\usepackage{isoent}    %\sfrac
%\usepackage{palatino} %font
\usepackage{newcent}   %font
\usepackage{sectsty} %custom section headings

\newcommand{\normalsections}{
\sectionfont{\noindent\rule{\textwidth}{0.015in}\\\nohang}
\subsectionfont{\noindent\rule{\textwidth}{0.005in}\\\nohang}
}

\newcommand{\columnsections}{
\sectionfont{\vspace*{-20pt}\noindent\rule{3.5in}{0.015in}\\\nohang}
\subsectionfont{\vspace*{-20pt}\noindent\rule{3.5in}{0.005in}\\\nohang}
}


\sectionfont{\noindent\rule{\textwidth}{0.015in}\\\nohang}
\subsectionfont{\noindent\rule{3.5in}{0.005in}\\\nohang}

\usepackage[Bjarne]{fncychap} %Canned chapter headings


\usepackage{multicol}
\usepackage[bookmarks=true,colorlinks,linkcolor=cyan,breaklinks]{hyperref}



%%%%Margins%%%
\topmargin 0pt
\advance \topmargin by -\headheight
\advance \topmargin by -\headsep
\textheight 8.9in
\oddsidemargin -0.25in
\evensidemargin \oddsidemargin
\textwidth 7in
\oddsidemargin -0.25in
%\setlength {\parindent} {0pt}



%%%Formatting%%%
\newcommand{\ability}[2]{\smallskip \noindent \textbf{#1} #2}
\newcommand{\shortability}[2]{\noindent\textbf{#1} #2\\}
\newcommand{\bolded}[1]{\noindent\textbf{#1}}
\newcommand{\itemability}[2]{\item \textbf{#1} #2}
\newcommand{\featname}[1]{\vspace*{0.1cm plus 0.2cm minus 0.05cm}\noindent\textbf{#1}\\}
\newcommand{\featnamelist}[1]{\vspace*{0.1cm plus 0.2cm minus 0.05cm}\noindent\textbf{#1}}

\newcommand{\descfeat}[2]{\featname{#1}\emph{#2}\\}
\newcommand{\business}[6]{
	\vspace{2pt}
	\noindent\hspace*{1cm}\textbf{#1}\\
	\noindent\hspace*{1cm}\textbf{Primary Skill:} #2\\
	\noindent\hspace*{1cm}\textbf{Secondary Skills:} #3\\
	\noindent\hspace*{1cm}\textbf{Capital:} #4\\
	\noindent\hspace*{1cm}\textbf{Resources:} #5\\
	\noindent\hspace*{1cm}\textbf{Risk:} #6
	\vspace{2pt}
}
\newcommand{\classname}[1]{\section{#1}}
\newcommand{\condition}[1]{\emph{#1}}
\newcommand{\quot}[1]{\emph{#1}\medskip}
\newcommand{\desc}[1]{#1 \medskip}
\newcommand{\example}[1]{\emph{#1}}
\newcommand{\magicitem}[1]{\emph{#1}}
\newcommand{\monster}[1]{\subsection{#1} \label{monster:#1}}
\newcommand{\monsterline}[2]{\textbf{#1:} #2\\}
\newcommand{\monstersizetype}[2]{\textbf{#1 #2}\\}
\newcommand{\spell}[1]{\emph{#1}}
\newcommand{\spelllist}[1]{\smallskip \noindent \underline{\textbf{#1}}}

% For the feats -- moved here from feats.tex because we added another 
% section(s) for feats. -Surgo
\newcommand{\minitabular}[1]{\begin{tabular}{p{0.25in}p{2.9in}} #1\\ \end{tabular}}
\newcommand{\babfeat}[7]{
	\noindent\minitabular{\multicolumn{2}{l}{\parbox{3in}{\textbf{#1}}}}
	\minitabular{\multicolumn{2}{l}{\parbox{3in}{\small #2}}}
	\minitabular{\raggedleft\textbf{\small \textbf{+0:}}& {\small #3}}
	\minitabular{\raggedleft\textbf{\small +1:} & {\small #4}}
	\minitabular{\raggedleft\textbf{\small +6:} & {\small #5}}
	\minitabular{\raggedleft\textbf{\small +11:} & {\small #6}}
	\minitabular{\raggedleft\textbf{\small +16:} &{\small #7}}
}
\newcommand{\skillfeat}[8]{
	\noindent\minitabular{\multicolumn{2}{l}{\parbox{3in}{\textbf{#1}}}}
	\minitabular{\multicolumn{2}{l}{\parbox{3in}{\small #2}}}
	\minitabular{\multicolumn{2}{l}{\parbox{3in}{\small\textbf{#3}}}}
	\minitabular{\raggedleft\textbf{\small \textbf{0:}} &{\small #4}}
	\minitabular{\raggedleft\textbf{\small 4:} &{\small #5}}
	\minitabular{\raggedleft\textbf{\small 9:} &{\small #6}}
	\minitabular{\raggedleft\textbf{\small 14:} &{\small #7}}
	\minitabular{\raggedleft\textbf{\small 19:} &{\small #8}}
}

\newcommand{\boxit}[1]{\frame{\parbox{\textwidth}{#1}}}

% A new box command -- the first argument is the box's header, the second the 
% contents of the box. This looks a lot better than \boxit. -Surgo
\newcommand{\abox}[2]{\vspace{5pt}
	\fbox{\begin{minipage}{0.8\linewidth}
	\setlength{\parindent}{0.15in}
	\textbf{#1}

	#2
\end{minipage}}
\vspace{5pt}}

\newcommand{\itemspace}{\setlength{\itemsep}{-1mm}\setlength{\topsep}{-1mm} }

\newcommand{\listone}{\begin{list}{$\bullet$}{\itemspace}}
\newcommand{\listprereq}{\begin{list}{\vspace*{-2pt}}{\itemspace}}
\newcommand{\listtwo}{\begin{list}{$\triangleright$}{\itemspace}}
\newcommand{\listthree}{\begin{list}{--}{\itemspace}}

% Bold the first argument in the listitem
\newcommand{\bolditem}[2]{\item \textbf{#1} #2}

\newcommand{\slashfrac}[2]{${}^{#1}$\hspace*{-1pt}\makebox[2pt]{$\diagup$}${}_{#2}$}
%\newcommand{\half}[0]{\slashfrac{1}{2}}
\newcommand{\half}[0]{\ensuremath{\sfrac{1}{2}} }
\newcommand{\third}[0]{\ensuremath{\sfrac{1}{3}} }
\newcommand{\fourth}[0]{\ensuremath{\sfrac{1}{4}} }
%\newcommand{\half}[0]{\ensuremath{\scriptscriptstyle 1/2}}
%\newcommand{\threefourths}[0]{\ensuremath{\sfrac{3}{4}}}


%\rule{\textwidth}{0.005in}

%\renewcommand{\sectionmark}[1]{\markright{\thesection\ \boldmath\emph{#1}\unboldmath}\\\rule{\textwidth}{0.005in}\\}

\begin{document}

%%%%%%% Front several pages
\input{cover}

\pagestyle{fancy}
%Fix the spacing so it's all reasonable
\linespread{.9}  \small  \normalsize \itemspace \normalsections

\tableofcontents

\chapter*{Foreword}

Unlike the Revised Necromancer Handbook, which is a compilation of the Necromancy rules as they stand, what you are reading now is the rules for Necromancy as they should be. We feel there is a need for this because despite (or let's not kid ourselves, because of) the considerable amount of space spent given over to Necromancy in officially sanctioned products, the classical Necromancer does not function under the rules as written. Vampires can't run or be staked, there aren't any prestige classes that make you any more of a necromancer than you are with the base classes, and honestly no one even knows how the basic necromancy spells work. Not because they are stupid, but because the rules for such things are contradictory in several key places.

\input{morality}

%%%%%%% Base Classes
\chapter{Necromancy with Class}

The degree to which the published classes of Necromancy aren't good causes people physical pain. The degree of malarkey that people are willing to attempt in order to use these classes in a half-way level appropriate way causes \emph{us} physical pain. While the flavor of published necromantic classes is frequently adequate or even engaging -- the mechanics just aren't there. You shouldn't have to cheat just to make a concept character \emph{nearly} the equal of a standard Cleric or Wizard. As a solution, we propose having actually mechanically viable prestige classes for the necromantically inclined to use:

\input{prestige/corpselight}
\input{prestige/uttercold}
\input{prestige/bonebladereaper}
\input{prestige/bonebladereaper_revised}
\input{prestige/skindancer}
\input{prestige/strangerwithburningeyes}
\input{prestige/masterofnecromanticmysteries}
\input{prestige/deathking}
\input{prestige/widowqueen}
\input{prestige/bonerider}
\input{prestige/thiefofsouls}
\input{prestige/soulmerchant}
\input{prestige/lurker}
\input{prestige/heartlessmage}
\input{prestige/speakerfordead}
\input{prestige/lorddamned}
\input{prestige/pumpkinking}

\chapter{Necromancers with Style}

\input{feats}

\chapter{The Necronomicon}

Necromancy as a school is possessed of some of the most powerful and game defining spells ever imagined in the worlds of Dungeons and Dragons. \spell{Magic jar}, \spell{wail of the banshee}, and \spell{clone} can practically be a world threatening plan for a BBEG all by themselves. In fact, that's been done several times. But Necromancy as a school suffers greatly for this attention. Though the earth shaking power for dark lords is well represented, the low levels of necromancy have been largely ignored by generations of authors. It is our intention to produce a short list of spells that allow a low level Necromancer to be memorable and effective without constantly falling back on the old stand-by of having Spell Focus: Conjuration.

\input{spells}

\chapter{New Rules}

The interaction of Undead with the rest of the rules is often less than satisfactory. Part of this is that the undead type itself is extremely overzealous in the game effects it provides. The fact that all undead don't need sleep means that vampires don't have to sleep in their coffins. The fact that undead don't have a Constitution score means that Ghouls can run for exactly zero rounds before they have to make a Con check (that they automatically fail) to continue (and also says it can ``run on indefinitely", a base contradiction that makes us sad). The fact that undead are immune to critical hits means that a vampire can't be staked through the heart (even if it was sleeping, which it isn't). But even beyond that fundamental error, the multitude of authors that compromise the Dungeons and Dragons design staff never seemed to get on the same page as to exactly what being undead means -- so a surprisingly large number of contradictory statements pepper the products.

And I'm not just talking about how they made an entire Deathless Type when there's already Ghosts (Alignment: Any) right in the core rules.

\input{undead}
\input{necromanticlocations}
\input{equipment}

\chapter{Monsters}

Perhaps the most important thing about necromancy is the undead creatures themselves.

\input{monsters}

\end{document}