\classname{Monster Rider}

\quot{Fuuwwuuuuuuuuuuuuhnhn!\\
See? Even Mishra knows that we can trample the Slaadi hordes today; isn't that right my pretty girl?\\
Fwweeehnuuhn!\\
That's a good girl. Now then you goblin cowards! Ready her armour and put those Tusk-Blades!\\
For Tonight, we dine on Frog!\\
-Kurzen Fiendson, Orcish Monster Rider}

\desc{Some warriors fight with axes, some use bows, others fight with cunning or astride a trusty steed.

A monster rider fights with their chosen fighting companion.}

\ability{Prerequisites:}{}
\listprereq
	\itemability{BAB:}{+2}
	\itemability{Skills:}{Ride 8 ranks}
	\itemability{Feats:}{Mounted Combat, Command or Leadership or any Necromantic Creation Feat}
\end{list}\vspace*{8pt}

\ability{Hit Die:}{d12}

\ability{Class Skills:}{The Monster Rider's class skills (and the key ability for each skill) are Balance (Dex), Climb (Str), Craft (Int), Diplomacy (Cha), Handle Animal (Cha), Heal (Wis), Hide (Dex), Intimidate (Cha), Jump (Str), Knowledge (Nature; The Planes, Religion, Arcana) (Int), Listen (Wis), Profession (Wis), Ride (Dex), Sense Motive (Wis), Spot (Wis), Survival (Wis), Tumble (Dex), and Use Rope (Dex).}

\ability{Skills/Level:}{6 + Int modifier}

\begin{table}[tbh]
\begin{small}
\begin{tabular}{lp{1.9cm}p{0.7cm}p{0.7cm}p{0.7cm}p{9cm}}
Level&Base Attack Bonus&Fort Save&Ref Save&Will Save&Special\\
1&+1&+2&+2&+2&Trusted Mount, Monster Rider Dice +1d6, Carrot and Stick, Mounted Attacks, Befriend Creatures, Monstrous Empathy\\
2&+2&+3&+3&+3&Monster Rider Ability, Speak with Monsters\\
3&+3&+3&+3&+3&Monster Rider Dice +2d6, Scarred hide\\
4&+4&+4&+4&+4&Monster Rider Ability, Genius Monster\\
5&+5&+4&+4&+4&Monster Rider Dice +3d6, Monster of Legend\\
\end{tabular}
\end{small}
\end{table}

\ability{Weapon and Armor Proficiency:}{The Monster Rider gains proficiency with the whip, composite shortbow, shortbow, lance, trident, all simple and martial polearms, simple thrown weapons and appropriate simple and martial weapons that have reach.}

\ability{Trusted Mount:}{As a full round action a Monster Rider may declare one helpful creature their Mount. This creature can gain any benefits that the Monster Rider class grants to their Mount.}

\desc{All of a Monster Rider's abilities that affect a mount last indefinately if the creature is the Monster Rider's cohort, or fade after one day if the creature is not the Monster Rider's chosen cohort, and such a creature's base CR cannot be increased; although they may be modified to gain new abilities.}

\desc{A Monster rider may only have one creature affected with the Trusted Mount ability.}

\ability{Carrot and Stick:}{There is no limit to how many tricks a Monster Rider can teach a creature, and their Handle Animal checks are not penalized for attempting to get a creature to perform a trick it does not know. In addition, a Monster Rider may teach tricks to any creature, even those without intelligence scores.}

\ability{Monster Rider Dice:}{Whenever a Monster Rider and their mount are sharing the same space, all BaB-derived attacks that either makes gain 1d6 more damage; this damage increases by 1d6 ever odd level.}

\ability{Befriend Creatures:}{A Monster Rider may make any creature up to Helpful with an animal empathy check by tending to any wounds that it may have suffered or offering it food. This check may be attempted once per day. Restraining a creature does not affect the Monster Rider's Animal Empathy check in any way, shape or form, and the Monster Rider may make as many attempts as they want with a creature, provided that they can heal the creature of wounds or offer it food.}

\ability{Mounted Attacks:}{A monster Rider counts any square of their mount as their own for purposes of reach for their weapon attacks (normally a rider would require a reach weapon to attack from the back of a creature much larger than itself).}

\ability{Monster Rider Ability:}{Choose an ability from the Monster Rider Abilities list.}

\ability{Speak with Monsters:}{Monster Riders can communicate with creatures as if they had a shared language. Creatures always seems to be able to catch the thrust of anything they say.}

\ability{Monstrous Empathy (Ex):}{As the Animal Empathy class ability of Druids and Rangers; class levels of the Monster Rider add to this check. This ability can be used on any creature that is non-humanoid in appearance, not just Animals or Magical Beasts.}

\ability{Scarred Hide:}{Your mount gains a bonus to its natural armour equal to 1/2 your character levels +2 and damage reduction equal to 1/2 your character levels +2 which is penetrated by nothing (a.k.a. DR X/-).}

\ability{Genius Monster:}{Your mount is affected as if an appropriate Maximized Awaken spell has been cast upon it. Creatures that are already intelligent, but were never affected by an Awaken effect in their current state gain the benefits of this effect as well. Use either the most effective or most appropriate Awaken Spell for the target; Awaken Undead for Undead mounts (with no maximum on intelligence) and Awaken Construct for Construct mouts; all other creatures should be affected as if Awaken (animals) had been cast on them.}

\ability{Monster of Legend:}{Your mount may choose to gain one of the following abilities, this ability cannot increase a Creature's CR about what it already is, and hit dice or the like may be removed in order to apply any of the following:}

\begin{list}{}{\itemspace}
\itemability{Hero of Monsters:}{Your mount may choose to remove its Racial Hit Dice and then increase its CR back to an appropriate value with class levels of an appropriate class.}
\item
\itemability{Planar Monster:}{Your mount may gain any inherited templates, they affect its CR normally, they do not affect anything else.}
\item
\itemability{Uneal Monster//Real Monster:}{Your mount may change its type from one to an other that it could theoretically be (Examples: Magical beast/animal/Vermin to -> Construct, Outsider, Undead. Dragon; Undead/Construct -> Magical Beast, Animal, Vermin), Recalculate CR and HD for this new type.}
\end{list}

\subsubsection{Monster Rider Abilities:}
\ability{Greater Ride(Ex):}{If you perform no other actions in your round may use a ride check in place of your mount's AC any time that they are attacked, and may use a Ride check in place of any Reflex saves that they would have to make. This ability places you outside the affected zone, and any movement made as a result of this effect is deducted from their movement in their next turn. If caught by an effect whose radius is larger than your mount could move in one round, then it simply is affected as if it had made the reflex save normally.}

\ability{Guided Mount Strikes (Ex):}{Any special class abilities that affect your attacks can be applied to any attacks that your mount makes with the same restrictions that your own class ability or abilities would have. Your character cannot make use of these class abilities in the same round that your mount makes use of them, nor can your mount use them if you have already used them.}

\ability{Monstrous Rage}{Your Mount can enter a Rage as the Races of War Barbarian can (upon being struck by an attack or upon hitting with a melee attack). Your mounts gains Rage Dice, Damage Reduction and Extra Movement as if they were a Barbarian whose level equals your Mount's Hit Dice. If your mount chooses to do Rage, your mount will attack targets in the following order of priority:
\listtwo
\item Any targets that it has observed attacking you
\item Any targets that have attacked it and have dealt it the most damage in this encounter
\item Any targets that have attacked it and damaged it
\item Any targets that have attacked it
\item The nearest target
\end{list}

\desc{This cannot be ended by your mount at-will, and can only be ended by a full round action on your part or if it falls unconsious.}

The Monster Rider may choose to direct where their mount moves and who they attack instead of allowing their mount to rage wildly, this is a standard action.}

\ability{Hulking Mount:}{Your Mount permanently increases by one size category, with all the appropriate changes. This ability may be taken more than once. It's effects stack.}

\ability{Follow Rider Anywhere:}{With a ritual taking one full day and supplies worth 100 gp per HD of your mount, your mount is able to decrease it's size by one category at will. In this smaller size it has the following modifiers applied -4 strength, +2 dex, +4 to hide, -4 on grapple checks and +1 to hit . Your mount may return to it's normal size or shrink as a swift action.}

This ritual can be performed more than once, the modifiers will stack. Mounts that have had their size reduced more than once may reduce their size as much or as little they want. If they choose to increase or decrease their size by one category at a time, this change takes a swift action, a mount that wishes to change their size by more than one size category takes a standard action to do so.