\section{Other Spheres}

These spheres are used by non-fiendish sphere users such as celestials and the Curator class.

\begin{multicols}{2}

\featname{\hypertarget{sphere:exorcism}{Exorcism}}
\textbf{Special: }{You gain the ability to turn undead three times per day as a good-aligned cleric of the same level as the class that granted you this sphere. You cannot use these turning attempts to power divine feats.}
\sphere
{Protection from Evil}
{Consecrate}
{Magic Circle against Evil}
{Dispel Evil}
{Disrupting Weapon}
{Banishment}
{Holy Word}
{Sunburst}
{Undeath's Eternal Foe (SpC)}
\spherecont{Freedom}

\featname{\hypertarget{sphere:mystery}{Mystery}}
\textbf{Special: }{You benefit from a constant undetectable alignment effect.}
\sphere
{Disguise Self}
{Silence}
{Blacklight}
{Greater Invisibility}
{Mirage Arcana}
{Forbiddance}
{Screen}
{Mind Blank}
{Etherealness}
\spherecont{Time Stop}


\featname{\hypertarget{sphere:piety}{Piety}}
\textbf{Special: }{If you succeed on a Sense Motive check (opposed by Bluff), you know the target's alignment. If you fail, you cannot make another attempt until you gain another rank of Sense Motive.}
\sphere
{Bless}
{Aid}
{Prayer}
{Lesser Globe of Invulnerability}
{Hallow}
{Heroes' Feast}
{Refuge}
{Holy Aura}
{Gate (Calling version only)}
\spherecont{Miracle}


\featname{\hypertarget{sphere:restraint}{Restraint}}
\textbf{Special: }{You gain a +3 bonus on saves against mind-affecting effects.}
\sphere
{Entangle}
{Web}
{Stinking Cloud}
{Black Tentacles}
{Wall of Stone}
{Repulsion}
{Forcecage}
{Prismatic Wall}
{Temporal Stasis}
\spherecont{Imprisonment}


\featname{\hypertarget{sphere:revelation}{Revelation}}
\textbf{Special: }{You receive a +2 bonus on all Knowledge checks, and never count as untrained in any Knowledge skill.}
\sphere
{Detect Evil}
{Detect Thoughts}
{Clairaudience/Clairvoyance}
{Divination}
{True Seeing}
{Find the Path}
{Legend Lore}
{Discern Location}
{Foresight}
\spherecont{Astral Projection}


\featname{\hypertarget{sphere:splendor}{Splendor}}
\textbf{Special: }{You gain a +2 bonus to Diplomacy, Intimidate, and Sense Motive.}
\sphere
{Charm Person}
{Eagle's Splendor}
{Flame Strike}
{Glorious Form (as Monstrous Form in the Dungeonomicon, except the new form appears glorious instead of monstrous)}
{Sympathy}
{Greater Heroism}
{Planar Ally}
{Sunburst}
{Weird}
\spherecont{Implosion}


\featname{\hypertarget{sphere:vigor}{Vigor}}
\textbf{Special: }{Whenever you heal someone with a spell-like ability, they recover additional HP equal to your ranks in the Heal skill.}
\sphere
{Cure Light Wounds}
{Shield Other}
{Cure Serious Wounds}
{Death Ward}
{Raise Dead}
{Heal}
{Greater Restoration}
{Greater Spell Immunity}
{Mass Heal}
\spherecont{True Resurrection}

\end{multicols}