\section{Necromantic Equipment}

\subsection{New Materials}

\ability{Boneblades:}{Boneblades are alchemically and necromantically hardened blades made from the bones of intelligent creatures, and the material can only be created by craftsmen with the Boneblade Master feat. For an unknown reason, they only retain their special properties if they are made into light slashing or piercing weapons.  Boneblades used in melee combat ignore the damage reduction of any undead creature and can hit incorporeal creatures as if they were magic weapons with the ghost touch property.

Boneblades made from dragon bones can be combined with the Dragoncrafter feat to produce items with both properties.

Cost: 1,000 gp per lb.}

\ability{Blood steel:}{Blood steel is steel that has been mixed with the blood of certain powerful creatures, making it redder than normal steel and with unusual properties. Weapons made of blood steel do 2 additional points of damage on a successful hit and any armor made of blood steel has an armor bonus two higher.  These effects only manifest if the blood steel is allowed to hungrily latch onto its user's flesh, at which point it reduces the bearer’s Constitution by two points.  Treat blood steel as normal steel for creatures with no Constitution or creatures protected from this effect.

Cost: 2,000 gp for a weapon, 1,000 gp for armor or shield.}

\ability{Black steel:}{Black steel is steel that has been mixed with necromantically charged obsidian, making it as sharp as adamantine and as dangerous as obsidian. Weapons made of black steel count as adamantine for all effects, but perform as if enhanced with the Ghost Touch and Wounding properties (without additional cost). Characters using items made of Black steel suffer one point of Wisdom drain for every day they are held, worn or carried.

Cost: 15,000 gp for a weapon, 24,000 gp for light armor, 28,000 gp for medium armor, 32,000 gp for heavy armor.}

\ability{Tainted obsidian:}{While normal obsidian often has interesting necromantic properties, tainted obsidian is an actual source of necromantic energy.  When a magic item has is made of tainted obsidian, it can provide uses or charges of its Necromancy effects in exchange for doing wisdom drain.

Cost: +10,000.}