\subsubsection{Robot Girl}
\quot{``Chiu...''}

\desc{While often constructed beings are nonliving in the strictest sense, sometimes a spellcaster or inventor will take it upon themself to build a being which is genuinely a living creature, albeit one which is composed of ceramic and metal rather than flesh and bone. Sometimes this is to create a servitor being which is less imposing than golems but more stable than simulacra, sometimes it's for other purposes.}

\desc{\listone
\item Type: Humanoid (Construct)
\item Ability Scores: no modifications
\item Medium: As Medium creatures, Robot Girls have no special bonuses or penalties due to their size.
\item Robot Girl base land speed is 30 ft.
\item Skills: +4 racial bonus to Search
\item Alignment: anything, really
\item Ghost in the Shell: As a creature with the Construct subtype, the Robot Girl has a Con score and everything, lacking Construct traits, but is treated as a Construct by any effect that specifically works differently against them, such as a Mace of Smiting or Favored Enemy (Constructs). Likewise, they can choose to count as a Construct for the purpose of taking feats, prestige classes and the like. Also, they have souls, and are partially made of metal, gaining a +1 Natural Armour Bonus.
\item Built: The Robot Girl can be healed by Repair spells as well as Cure spells, and gains a +4 bonus against Poison and Disease. They need not sleep, but can do so if they want. They still need food.
\item Lightning-Powered: They gain Electricity Resistance 10 + Hit Dice. Any time any Electricity is resisted, they count as having eaten a full meal. It is delicious.
\item Laser Eyes (Su): Can be used once per two rounds, with a 30' range. Requires a Ranged Touch Attack, and deals 1d4 Light damage per hit die.
\item Scanner (Sp): You may cast Detect (Magic/Poison/Evil/Good/Chaos/Law) at will.
\end{list}}

\descfeat{Steel Angel}{You are extra robotic but extra sweet}
\textbf{Requirement:}{Robot Girl}

\ability{Benefit:}{You gain Damage Reduction/Adamantine equal to your hit dice, and any [Charm] effects you use gain a +1 Racial bonus to the save DC.}

\ability{5 HD:}{Once per day you can fire a salvo of missiles from your shoulders/wrists/knees/boobs as a (Su) ability, out to 150'. They explode in a 30' burst, dealing 1d6 Fire damage per HD and 1d6 Sonic damage per HD, Ref half (Con-based).}

\ability{10 HD:}{You gain a Fly speed of 30' (Good) as you jet around the place.}

\ability{15 HD:}{Once per day you may fire a Photon Beam from your face/chest as a (Su) ability. This extends in a 100' long line, hitting everyone in the area with a Disintegrate effect. The save DC is Con-based.}