\classname{Ninja}
\vspace{-8pt}
\quot{"Nin Nin!"}

\ability{Playing a Ninja:}{Ninja get to add Int to a lot of things and have Save DCs based on Int, so it benefits them to make it one of their better scores, but quite so radically as a Wizard. They also benefit from a high Dex, with other ability scores chosen as your style dictates.}

\ability{Alignment:}{Ninja can be loyal to their clan master or masterless brigands, if you think that describes Law or Chaos (similarly they can be wet or dry, if you use that metric). They can be murderous assassins or protectors of the people. Ninja can be of any alignment.}

\ability{Races:}{anyone can be a Ninja, though humans and elfs gravitate towards it, as do halflings. There are only a few Mountain Giant Ninja out there; no more than a dozen, tops. Kappa are weirdly keen on being Ninja.}

\ability{Starting Gold:}{4d4*10 gp (100 gp)}

\ability{Starting Age:}{As Rogue}

\ability{Hit Die:}{d6}

\ability{Skills/Level:}{6 + Int Bonus + special*}

\ability{Class Skills:}{The ninja's class skills (and the key ability for each skill) are Appraise (Int), Balance (Dex), Bluff* (Cha), Climb (Str), Diplomancy (Cha), Disable Device (Int), Disguise* (Cha), Escape Artist* (Dex), Forgery* (Int), Gather Information (Cha), Hide* (Dex), Intimidate (Cha), Jump* (Str), Knowledge (Any) (Int), Listen (Wis), Move Silently* (Dex), Perform (Dance/Stage Theatre) (Cha), Profession (Any), Ride (Dex), Search* (Int), Sense Motive (Wis), Spot (Wis), Survival (Wis), Swim (Str), Tumble* (Dex), and Use Rope (Dex)}
\emph{*Select one of these skills upon taking your first Ninja level. The Ninja always has maximum ranks in this skill without spending skill points.}


\begin{table}[tbh]
\begin{small}
\begin{tabular}{lp{3cm}p{0.7cm}p{0.7cm}p{0.7cm}p{7cm}p{0.7cm}l}
Level  &Base Attack Bonus &Fort Save &Ref Save &Will Save &Special & Jutsu Points &AC Bonus\\
1st    &+1              &+0 &+2  &+2  & Armoured in Life, Surprise!, Jutsu Use, Too Fast By Half	&1 &+4\\
2nd    &+2              &+0 &+3  &+3  & Wall Jump, Ninja'd &2 &+5\\
3rd    &+3              &+1 &+3  &+3  & Invisible, Speed of Thought  &2 &+5\\
4th    &+4              &+1 &+4  &+4  & Dark Stalker, Hypnotic Gaze  &3 &+6\\
5th    &+5              &+1 &+4  &+4  & Master of Disguise, Mystic Art  &3 &+6\\
6th    &+6/+1           &+2 &+5  &+5  & Double-Jump, Scuttler  &3 &+7\\
7th    &+7/+2           &+2 &+5  &+5  & Anti-Noise Aura  &4 &+7\\
8th    &+8/+3           &+2 &+6  &+6  & Log Trick, Go Ninja Go Ninja Go  &4 &+8\\
9th    &+9/+4           &+3 &+6  &+6  & Master of Escape, Lightning Strike  &4 &+8\\
10th   &+10/+5          &+3 &+7  &+7  & Mystic Art, Close Combat  &4 &+9\\
11th   &+11/+6/+6       &+3 &+7  &+7  & Shadow Arts I: Veils of Shades &5 &+9\\
12th   &+12/+7/+7       &+4 &+8  &+8  & Too Fast By \textthreequarters &5 &+10\\
13th   &+13/+8/+8       &+4 &+8  &+8  & Spirit Blade &5 &+10\\
14th   &+14/+9/+9       &+4 &+9  &+9  & Perfect Disguise &5 &+11\\
15th   &+15/+10/+10     &+5 &+9  &+9  & Mystic Art, Perpetual Distraction &5 &+11\\
16th   &+16/+11/+11/+11 &+5 &+10 &+10 & Opportunistic Strike &6 &+12\\
17th   &+17/+12/+12/+12 &+5 &+10 &+10 & Murder Punch &6 &+12\\
18th   &+18/+13/+13/+13 &+6 &+11 &+11 & Hidden from Magic&6 &+13\\
19th   &+19/+14/+14/+14 &+6 &+11 &+11 & Shadow Arts II: Nightmare Realm	6 &6 &+13\\
20th   &+20/+15/+15/+15 &+6 &+12 &+12 & Mystic Art, Fully Too Fast, Clan Master &6 &+14\\
\end{tabular}
\end{small}
\end{table}

\smallskip\noindent All of the following are class features of the ninja.

\ability{Weapon and Armor Proficiency:}{the Ninja is proficient with all Simple Weapons, all Martial Weapons, and two Exotic Weapons of the player's choice. The Ninja is not proficient with any armour.}

\ability{Special:}{If the Ninja is a Kappa or some other variety of turtle-person, he is instead only proficient with the Quarterstaff, Sai, Long Sword, Nunchaku and Shuriken, but gains Two Weapon Fighting as a Bonus Feat.}

\ability{A Note on Standard Attacks:}{various abilities require a "Standard Attack" be made. This is a regular attack made as a Standard action.}

\ability{Jutsu Use (Su):}{as she gains levels, the Ninja gains a number of Jutsu Points. With a Free action, these Jutsu Points can be assigned to different things, but once assigned, they are stuck that way until a minute is spent meditating, which reallocates all points to "unspent". Jutsu Points can be allocated to any of the following, adding her Int modifier as an Insight bonus:\smallskip

\begin{itemize}\itemspace
    \item Attack rolls
    \item Damage rolls
    \item Saving throws
    \item Spell resistance (must possess it already)
    \item Fast Healing (must possess it already)
    \item Dex-based Skill checks
    \item Cha-based skill checks
    \item Wis-based skill checks
    \item Str-based skill checks
    \item Constitution Score
    \item All pre-existing Energy Resistances (must already possess at least one type)
    \item Total Damage Reduction (in any instance, after adding various types. If none apply, treat this alone as /-)
\end{itemize}

\smallskip Alternately, a Jutsu Point can be "spent" (unusable until renewed through the minute of meditation) to regain a daily use of a Spell-like Ability gained from the Ninja Class, to activate certain abilities, or to re-roll Initiative. A Jutsu point spent in such a way cannot have already been allocated - it must be left in reserve and then spent.}

\ability{Armoured In Life (Su):}{As long as she is not using armour or a shield that he is not proficient with, the Ninja gains a special Armour bonus. This Armour bonus even applies against Touch Attacks and Incorporeal Touch Attacks, and has a value of +4. Every two levels, it increases by +1. If wearing armour with which she is proficient, she may add its Enhancement bonus to this.\smallskip

If the Ninja has levels of Monk, then Ninja levels count as Monk levels for the purpose of determining the size of the bonus.}

\ability{Surprise! (Ex):}{The Ninja is good at taking advantage of unprepared foes. She has the Edge against any enemy who is denied their Dexterity bonus to Armour Class, and once per turn may make an Attack of Opportunity against one such foe.}

\ability{Too Fast By Half (Ex):}{The Ninja multiplies her Base Speed by 1.5 whenever she is not using armour or a shield with which she is not proficient.}

\ability{Wall Jump (Ex):}{At second level, the Ninja can use her jumping skills to bounce off walls, and to slide down them as well. This prevents her from suffering falling damage when adjacent to a wall, and allows her to halve any Jump DC if she starts the jump adjacent to a wall. If effectively flanked by two adjacent walls, she gains a Climb speed equal to her land speed.}

\ability{Ninja'd (Ex):}{Ninja always seize the initiative at the last moment, before others can react. At second level, if there is no Surprise Round for a given combat, the Ninja gains her own Surprise Round. If there is a Surprise Round, the Ninja automatically is able to act in it. The Ninja can only access this ability if she possesses more levels of Ninja than any other class.}

\ability{Invisible (Su):}{The third level Ninja may become Invisible at will, as per the spell \spell{invisibility}. This still requires a Standard Action to activate and ends if she attacks.}

\ability{Speed of Thought (Ex):}{The third level Ninja moves so fast she gets speed lines when she runs. She may add her Intelligence bonus to her Initiative, and to all Trip, Bullrush, Disarm and Grapple attempts against foes who are denied their Dexterity bonus to Armour Class.}

\ability{Dark Stalker (Ex):}{At level four, the Ninja can see and hear Incorporeal and Ethereal creatures normally, as if they lacked those traits, so they must make Hide and Move Silently checks if they want to sneak up on her.}

\ability{Hypnotic Gaze (Su):}{The level four Ninja gains a Gaze attack out to 30' that causes targets who fail a Will Save (Intelligence-based) to become Fascinated until attacked or until they lose sight of the Ninja. By spending a Jutsu Point she can add a Suggestion effect to a creature affected by the Gaze, though they are allowed a separate save against this effect (same DC). A creature who successfully saves is immune to that Ninja's Gaze for the rest of the day.}

\ability{Master of Disguise (Su):}{With a DC 15 Disguise check, a fifth-level Ninja can replicate a Disguise Self effect. With a DC 20 Disguise check, she can replicate an Alter Self effect.}

\ability{Mystic Art:}{At level five, and every five levels thereafter, the Ninja gains a special Mystic Art that can be selected from the following list. Some have minimum level requirements:}
\begin{itemize}\itemspace\begin{small}
    \item{\ability{Vital Strike (Ex):}{once per round, when the Ninja strikes a foe who is denied their Dexterity bonus to Armour Class, the foe must make a Fortitude Save (Intelligence-based) or be Nauseated for 1 round.}}
    \item{\ability{\spell{Fog Cloud} (Sp):}{each use of this ability requires spending a Jutsu point.}}
    \item{\ability{\spell{Hold Monster} (Sp):}{each use of this ability requires spending a Jutsu point. Furthermore, passing the save still results in suffering 5d6 non-lethal damage and taking a -3 penalty to Will saves for 1 round. The Save DC is Intelligence-based.}}
    \item{\ability{Kagenui (Ex):}{with a Standard attack, the Ninja forces a successfully hit opponent to attempt a Fort Save (Intelligence-based) or be Anchored in place for 3 rounds.}}
    \item{\ability{Mafuujin (Su):}{with a Standard attack, the Ninja forces a successfully hit opponent to attempt a Will Save (Intelligence-based) or be Confused for 1 minute. She may elect for this to deal non-lethal damage at no penalty.}}
    \item{\ability{Flight (Su):}{this requires a Jutsu Point be assigned to flying - doing so grants a Fly speed of 30' (Average), +5' per level. Minimum level 10.}}
    \item{\ability{Not Visible (Su):}{this requires a Jutsu Point be assigned to not being visible - doing so grants Improved Invisibility that cannot be beaten by True Seeing or See Invisible, only by Invisibility Purge - although attacking makes her partially visible as normal, invisibility resumes at the beginning of her next turn. Minimum level 10.}}
    \item{\ability{Magic Focus (Su):}{by spending a Jutsu Point, the Ninja regains a Daily or Hourly use of a Spell-like Ability from any source, even magic items. Minimum level 10.}}
    \item{\ability{Shini Kaze (Ex):}{with a Standard attack, the Ninja forces a successfully hit opponent to attempt a Fort Save (Intelligence-based) or be Poisoned (2d6 Con/2d6 Con). Minimum level 10.}}
    \item{\ability{Tougenkyo (Su):}{with a Standard attack, the Ninja forces a successfully hit opponent to attempt a Fort Save (Intelligence-based) or be Blinded and Cursed (ability score penalty) for 1 minute. Minimum level 10.}}
    \item{\ability{Devil Ghost Body (Su):}{a Jutsu point may be assigned to Armour Class, adding the Ninja's Intelligence bonus as an Insight bonus to it. Minimum level 10.}}
    \item{\ability{Spell Stowaway (Su):}{any time someone within 30' casts a spell on themselves or on a target within 30' of the Ninja, the Ninja, after attempting to identify the spell, may choose to also be affected by it, spending an Immediate action to do so. Minimum level 15.}}
    \item{\ability{Disregard Me (Su):}{this requires a Jutsu Point be assigned to it, resulting in the Ninja being impossible to detect with Divinations unless she wants to be detected by them, and she never provokes Attacks of Opportunity while benefiting from it. Enemies can't even use readied actions against her. However abilities that allow Attacks of Opportunity to be made even against people who don't provoke, or that force people to provoke, still work. Minimum level 15.}}
    \item{\ability{Min Min Min (Su):}{with a Standard attack, the Ninja forces a successfully hit opponent to attempt a Fort Save (Intelligence-based) or fall Asleep for 10 minutes. Minimum level 15.}}
    \item{\ability{You Can't See Me (Su):}{this requires a Jutsu Point to be assigned to it. The Ninja may use an Immediate action to cause an attack, spell or effect directed at her to instead target the attacker/originator of the effect. They must continue to use it at full force, with the full bonus they would have used against him, and cannot attempt to redirect the effect or force it to miss or fail. Minimum level 20.}}
    \item{\ability{Forbidden Technique (Su):}{by spending a Jutsu Point, the Ninja can use a Full Round action to force a foe within 30' to attempt a Will Save (Intelligence-based). If they pass, they suffer 2d6 negative levels. If they fail, they are instantly annihilated, and can only be brought back by journeying to the plain their soul is sent to and performing a great quest. Minimum level 20.}}
    \item{\ability{Harem No Jutsu (Su):}{by spending a Jutsu Point, the Ninja bestows a Hold Monster effect on all targets within 50 feet. The save is Intelligence-based, and those who are held three rounds in a row must make a Fortitude Save (same DC) or die, but die happy. Those who pass the initial save are Stunned for 1 round and suffer 5d6 non-lethal damage. Minimum level 20.}}
    \item{\ability{Yukkikage (Ex):}{by spending a Jutsu Point, the Ninja may use one Full Action to get two Full Actions, though these may not further be doubled by this ability. Minimum level 20.}}
\end{small}\end{itemize}

\ability{Double-Jump (Ex):}{The sixth-level Ninja is able to jump while already in the air. The height and length of any Jump she makes may be doubled if she wishes, and she may turn corners and other such things while jumping.}

\ability{Scuttler (Ex):}{The level six Ninja is great at moving about in all situations. She may move at regular speed while being stealthy, and does not suffer penalties for difficult terrain.}

\ability{Anti-Noise Aura (Su):}{The level seven Ninja is so stealthy that she passes it on to her allies: at will she may radiate a 20' radius Aura of Silence, masking all noises made inside it as though inside a sound-proof dome.}

\ability{Log Trick (Ex):}{Just when you think you have a Ninja, you actually have a log. The Ninja is behind you. At level eight, the Ninja may Dimension Door up to 30' away with an Immediate action, leaving a Mirror Image behind for 4 rounds (or until interacted with). When the image ends, a log appears in its place. By spending a Jutsu Point, she may perform this as an Immediate Action. This ability cannot be performed in armour with which she is not proficient.}

\ability{Go Ninja Go Ninja Go (Ex):}{The eighth-level Ninja may make a Full Round action (or a Standard action and a Move action, or two Move actions) and a Swift action in the Surprise round, as opposed to just one Partial action. This does not apply when wearing armour with which she is not proficient.}

\ability{Master of Escape (Ex):}{The ninth-level Ninja is hard to pin down. Any Escape Artist check may be made as a Standard action that does not provoke, and she may make Epic checks at the following DCs: Extremely Tight Space (a chimney, a glass jar, someone's anus) DC 30, Walls of Force DC 40 or Caster Level + 5, whichever is higher. She also adds her Intelligence Bonus to all attempts to escape a Grapple. None of this applies when wearing armour with which she is not proficient.}

\ability{Lightning Strike (Ex):}{When flying or jumping, the level nine Ninja may make a special charge attack. She moves in a straight line, up to twice her Movement Speed, and makes one attack roll, applying it against everyone in this line. She gains the charge bonus and penalty still, and anyone successfully hit must pass a Fortitude Save (Intelligence-based) or be Staggered for one round. This ability cannot be used while wearing armour with which she is not proficient, and performing it leaves her Staggered on her next turn.}

\ability{Close Combat (Ex):}{At level ten, the Ninja functions just fine when in a Grapple - she can act normally, even moving - even if Pinned - and is not denied her Dexterity bonus to Armour Class. Furthermore, her opponent is denied their Dexterity bonus to Armour Class against her attacks.}

\ability{Shadow Arts I: Veils of Shades (Sp)}{At level eleven, the Ninja gains the following Spell-like abilities: Hallucinatory Terrain once per day, Darkness at will, and Mirror Image a number of times per day equal to her Intelligence bonus (minimum 1).}

\ability{Too Fast By \textthreequarters (Ex):}{At level twelve, the Ninja becomes even faster. She may make a Move action as an Immediate action, and makes 10' steps of adjustment instead of 5' steps (unless she only wants to move 5'). With a DC 25 Balance check, she can instead make 15' steps of adjustment.}

\ability{Spirit Blade (Su):}{The thirteenth-level Ninja can even cut ghosts. All of her attacks affect Incorporeal and Ethereal creatures as though they lacked those abilities, and ignore Armour and Shield bonuses to Armour Class (but not Natural Armour), as well as all kinds of Damage Reduction.}

\ability{Perfect Disguise (Su):}{At level fourteen, the Ninja can use a DC 35 Disguise check and ten minutes to disguise herself almost perfectly into any creature - she can appear as a creature of any Type, Subtypes or Size, providing the creature does indeed appear at all. This simply results in her looking like the creature, not taking on any of its abilities, and taking no penalties to future Disguise or Bluff checks to blend into Storm Giant society or the Pixie court or whatever. True Seeing does not trump this, though effects that force creatures to revert to their true form will work.}

\ability{Perpetual Distraction (Su):}{People are always distracted when fighting Ninjas. At level fifteen, the Ninja always counts as flanking any enemy she attacks.}

\ability{Opportunistic Strike (Ex):}{At level sixteen, the Ninja is really good at taking advantage of foes who are already hobbled. Falling Prone provokes an Attack of Opportunity from the Ninja, and Prone foes are denied their Dexterity bonus to Armour Class against her.}

\ability{Murder Punch (Ex):}{People die when they are killed. All attacks made by a seventeenth-level Ninja ignore Regeneration - the damage is just there until healed (naturally or by magic), and is very much real damage. Anyone slain by her can only be brought back by a Wish, Miracle, True Resurrection or great journey to the afterlife and personally defeating the setting's god of death.}

\ability{Hidden From Magic (Su):}{At eighteenth level, the Ninja may, after attempting to identify a spell being cast from within 30', use an Immediate action to be excluded from the Targets or Area of Effect. She may even use this to walk unimpeded through a spell that has already been cast such as a Wall of Fire. A weapon enhanced by a spell will still attack her just fine - the target is the weapon, not her - likewise, a non-magical object conjured by magic (such as a lump of rock) affects her normally after the initial conjuration, only providing potential immunity to the effects of it being summoned around her in the first place.}

\ability{Shadow Arts II: Nightmare Realm (Sp):}{At level nineteen, the Ninja may cast Nightmare Terrain three times per day. Every round, one target of his choice within the Nightmare Terrain area can be targeted with a Shadow Well effect (Intelligence-based). She may also once per day transform into a Nightwalker as a Spell-Like Ability, completely replacing her character sheet with the monster entry, save for her Alignment and the ability to transform back as a Free Action, or automatically if destroyed or after one hour. No effects - damage, ongoing spells, magic items, class features, diseases, anything - carry from one form to the other, though her condition is "frozen" in place when she transforms: if she is poisoned, the transformation won't save her from saving against secondary damage, it just takes place a lot later.}

\ability{Clan Master (Ex):}{At level twenty, the Ninja becomes Ageless and Immortal. She needn't eat, drink or sleep, and will never die of old age, appearing to eternally be in the prime of her youth. Bonuses and penalties from age no longer apply, and never will apply.}

\ability{Fully Too Fast (Ex):}{The twentieth-level Ninja's speed improves from one and a half times their base speed to double the base speed. Additionally, she may \spell{Teleport} at will, that's how fast she moves.}