\classname{Warmage}
\vspace{-8pt}
\quot{The biggest armies require the most resources, this is why I require a mountain of gold.}

\desc{Wizards in general are pretty good on the battlefield. When it comes to killing lots of people at a time, casters excel at that if they feel like it. Not that a single great warrior can't hack through an army by the time everyone else has done their shoes up, but it's always the explosions of flame that people remember.}

\desc{But some casters are specifically designed for the battlefield. Not only do they call ordinance down, annihilating large numbers of people at a time, but they can also take care of the stuff that only the good commanders think of: food, drink, digging trenches, building fortifications, gathering intel, making the battlefield conditions perfect. When you need someone like that, you turn to the Warmage.}

\ability{Starting Gold:}{5d4x10 gp (125 Gold)}

\ability{Starting Age:}{As Wizard}

\ability{Hit Die:}{d6}

\ability{Class Skills:}{The Warmage's class skills (and the key ability for each skill) are Appraise (Int), Bluff (Cha), Concentration(Wis), Craft (Int), Diplomacy(Cha), Gather Information (), Intimidate (Cha), Listen (Wis), Knowledge (Any) (Int), Ride (Dex), Spellcraft (), Spot (Wis), and Survival (Wis).}

\ability{Skills/Level:}{4 + Int modifier.}

\begin{table}[htb]
\begin{small}
\begin{tabular}{lp{3cm}p{0.7cm}p{0.7cm}p{0.7cm}l}
Level  &Base Attack Bonus &Fort Save &Ref Save &Will Save &Special\\
1st &+0 &+0 &+2 &+2 &Elemental Exchange (Fire and Cold), Spellcasting\\
2nd &+1 &+0 &+3 &+3 &Purify Food and Drink\\
3rd &+1 &+1 &+3 &+3 &Create Food and Water\\
4th &+2 &+1 &+4 &+4 &Elemental Exchange (Acid, Electricity)\\
5th &+2 &+1 &+4 &+4 &More Targets\\
6th &+3 &+2 &+5 &+5 &Limitless Spell Force\\
7th &+3 &+2 &+5 &+5 &Explosive Spells\\
8th &+4 &+2 &+6 &+6 &Elemental Exchange (Sonic)\\
9th &+4 &+3 &+6 &+6 &Weaken Defences\\
10th &+5 &+3 &+7 &+7 &Even More Targets\\
11th &+5 &+3 &+7 &+7 &Rallying Spells\\
12th &+6/+1 &+4 &+8 &+8 &Elemental Exchange (Force, Negative)\\
13th &+6/+1 &+4 &+8 &+8 &Heroes Feast\\
14th &+7/+2 &+4 &+9 &+9 &Instant Fortress\\
15th &+7/+2 &+5 &+9 &+9 &Improved Targeting\\
16th &+8/+3 &+5 &+10 &+10 &Energy Exchange (Positive)\\
17th &+8/+3 &+5 &+10 &+10 &Dimensional Fortress\\
18th &+9/+4 &+6 &+11 &+11 &Unbreakable Barricades\\
19th &+9/+4 &+6 &+11 &+11 &Craft Arcane Ordinance\\
20th &+10/+5 &+6 &+12 &+12 &Arcane Exchange (Vile), All The Targets\\
\end{tabular}
\end{small}
\end{table}

\begin{table}[htb]
\begin{small}
\noindent\begin{tabular}{lllllllllll}
 & \multicolumn{10}{c}{Warmage Spells Per Day}\\
 &0 &1 &2 &3 &4 &5 &6 &7 &8 &9\\
1 &4 &2 &- &- &- &- &- &- &- &-\\
2 &5 &3 &- &- &- &- &- &- &- &-\\
3 &6 &4 &2 &- &- &- &- &- &- &-\\
4 &6 &5 &3 &- &- &- &- &- &- &-\\
5 &6 &6 &4 &2 &- &- &- &- &- &-\\
6 &6 &6 &5 &3 &- &- &- &- &- &-\\
7 &6 &6 &6 &4 &2 &- &- &- &- &-\\
8 &6 &6 &6 &5 &3 &- &- &- &- &-\\
9 &6 &6 &6 &6 &4 &2 &- &- &- &-\\
10 &6 &6 &6 &6 &5 &3 &- &- &- &-\\
11 &6 &6 &6 &6 &6 &4 &2 &- &- &-\\
12 &6 &6 &6 &6 &6 &5 &3 &- &- &-\\
13 &6 &6 &6 &6 &6 &6 &4 &2 &- &-\\
14 &6 &6 &6 &6 &6 &6 &5 &3 &- &-\\
15 &6 &6 &6 &6 &6 &6 &6 &4 &2 &-\\
16 &6 &6 &6 &6 &6 &6 &6 &5 &3 &-\\
17 &6 &6 &6 &6 &6 &6 &6 &6 &4 &2\\
18 &6 &6 &6 &6 &6 &6 &6 &6 &5 &3\\
19 &6 &6 &6 &6 &6 &6 &6 &6 &6 &4\\
20 &6 &6 &6 &6 &6 &6 &6 &6 &6 &5\\
\end{tabular}
\end{small}
\end{table}

\smallskip\noindent All of the following are Class Features of the Warmage class.

\ability{Weapon and Armor Proficiency:}{Warmages are proficient with all Simple and Martial weapons. Warmages are proficient with Light and Medium Armor but not with shields of any kind. The Warmage does not suffer Arcane Spell Failure for armour with which he is proficient.}

\ability{Spellcasting:}{The Warmage automatically knows every spell on their class list that they are a high enough level to cast. They cast spells spontaneously, without preparation. The ability score that determines their spellcasting is Intelligence. Unless stated otherwise, any Warmage class features that affect her spells will only affect those gained from Warmage levels, not other classes.}

\ability{Elemental Exchange (Su):}{as the Warmage gains levels, he gains mastery over the elements. Any time he casts a spell that deals [Energy] damage, he may change the type of energy to another.
At first level, he can only change it to Fire or Cold damage.
At level four, he may also change it to Acid or Electricity.
At level eight, he may change it to Sonic damage.
At level twelve, he may change it to Force or Negative Energy. Negative heals Undead.
At level sixteen, he may change it to Positive Energy, which heals most creatures.
At level twenty, he may also make the damage Vile (such as "Vile Fire Damage"), should anyone survive.
When the damage is changed, the descriptor is as well - for example, a Cold Fireball is a [Cold] effect.}

\ability{Purify Food and Drink (Sp):}{The second-level Warmage can cast \spell{Purify Food and Drink} at will}

\ability{Create Food and Water (Sp):}{Starting at third level, the Warmage can conjure food and water at will. Up to one gallon of fresh, clean water per level and one pound of hard tack per level is conjured.
\smallskip\noindent
This is an Instantaneous Conjuration [Creation] effect, and is effectively a second-level spell.}

\ability{More Targets (Su):}{at level five, the Warmage gets better at affecting groups of people. Any of his spells that specify multiple targets now affect twice as many. Those that affect a number of hit dice worth of creatures affect twice as many hit dice of creatures. Spells with an area of effect see that area of effect doubled.}

\ability{Limitless Spell Force (Su):}{Starting at level six, there is no upper limit on how far the spells of the Warmage scale. For instance, Fireball will do 1d6 damage per level even beyond the 10d6 at level 10, and Scorching Ray will create 5 rays at level 19.}

\ability{Explosive Spell (Su):}{at level seven, the Warmage may make a spell Explosive, however doing so prevents it from having More Targets (or a better variant). An Explosive Spell must have an area of effect, and any target that takes damage must make a Fortitude Save (DC 10 + half his level + his Intelligence Bonus). If the Fortitude Save is failed, they are flung out of the area, and half the distance travelled again, in the direction from which they were hit. For instance, if 10' away from the edge of a Fireball, they would move 10 + 5 = 15' away. If caught at the start of a 120' long Line, they would travel 180'. At the end of this movement, they land Prone.}

\ability{Weaken Defences (Su):}{Whenever a target takes damage from a spell cast by the ninth-level Warmage, their defences are lowered. Until the beginning of the Warmage's next turn, they take a penalty to their Armour Class, Damage Reduction, Spell Resistance and all Energy Resistances. The penalty equals the Spell Level, and can't reduce the trait below zero.}

\ability{Even More Targets (Su):}{At level ten, when "More Targets" applies, the number is tripled instead of doubled. The Warmage affects a lot of targets now.}

\ability{Rallying Spell (Su):}{At level eleven, the Warmage may make a spell Rallying, though doing so prevents it from having More Targets (or a better variant). Nor can it be Explosive in order to be Rallying. There is a large blast of light and sound for one round, and all allies within line of sight remove all [Fear] effects. Additionally, they gain a Morale Bonus on Attack and Damage rolls and Will Saves equal to the Spell Level for one round.}

\ability{Heroes Feast (Sp):}{Once per day at level thirteen, the Warmage may cast \spell{Heroes' Feast} as a Spell-Like Ability. Despite not actually being a spell, it feeds and affects additional people, gaining the benefits of More Targets (and the variants).}

\ability{Instant Fortress (Sp):}{At level fourteen, the Warmage may summon an Instant Fortress once per day. This requires one minute of concentration. The fortress so created functions as the item of the same name, but when packed up, all damage is recovered instantly for the next day. If destroyed, it takes a full week for another to be called forth. By extending the concentration time to ten minutes, it creates two towers, connected by a wall up to 50' long, 25' high and 10' thick, with battlements and arrow slits. This is a seventh-level Conjuration [Creation] effect.}

\ability{Improved Targeting (Su):}{At level fifteen, when the Warmage casts a spell with an area of effect, he may designate up to one creature or 5' square per hit die to be exempt from this. The exemptions are not multiplied by More Targets.}

\ability{Dimensional Fortress (Sp):}{Once per day, the seventeenth-level Warmage can cast Mordenkainen's Magnificent Mansion. Anybody who attempts to gain entry uninvited triggers an alarm and a Disintegrate effect with a Save DC of 10 + half his hit dice + his Intelligence Bonus.}

\ability{Unbreakable Barricades (Sp):}{With a Full Round Action, the eighteenth-level Warmage can affect a square mile worth of walls and other structures, reinforcing them. This is a Permanent, Dismissable effect, giving the structures the Hardness of Adamantium and treating them with Force Effects - any damage must first bypass the Force Effects. The Force effects also mean Ethereal and Incorporeal creatures cannot move through the barriers, and the effect also blocks any Teleportation. This is considered to be a ninth-level Abjuration [Force] effect.}

\ability{Craft Arcane Ordinance (Su):}{At level nineteen, the Warmage gains the ability to create magic weapons: war machines. This uses whatever magic crafting rules you are using. He may also craft Constructs, ships and the like, but not personal arms and armour.}

\ability{All the Targets (Su):}{The level twenty Warmage ignores planar boundaries when casting spells - he can just have his spells reach into the Ethereal plane, for instance.}

\spelllist{Warmage Spell List:}
\small
    \ability{Cantrips:}{\emph{Dancing Lights, Darkness, Detect Magic, Detect Poison, Disrupt Undead, Ghost Sound, Light, Magic Missile, Mending, Message, Purify Food and Drink, Rouse, Stand}}
    
    \ability{Level 1:}{\emph{Alarm, Detect Scrying, Detect Secret Doors, Disguise Self, Endure Elements, Entangle, Firespray, Grease, Mass Rouse, Mass Stand, Mount, Obscuring Mist, Shocking Grasp, Silent Image, Ventriloquism}}
    
    \ability{Level 2:}{\emph{Arcane Lock, Continual Flame, Fireball, Jet of Steam, Gust of Wind, Incendiary Slime, Locate Object, Mass Enlarge Person, Mass Reduce Person, Misdirection, Protection From Arrows, Pyrotechnics, Resist Energy, Scare, Scorching Ray, See Invisibility, Whispering Wind}}
    
    \ability{Level 3:}{\emph{Caustic Mire, Caustic Smoke, Deep Slumber, Dispel Magic, Flame Arrow, Fly, Horrid Sickness, Invisibility Sphere, Lightning Bolt, Major Image, Mass Resist Energy, Mordenkainen's Faithful Hound, Nondetection, Protection From Energy, Rage, Sleet Storm, Stinking Cloud, Wall of Fire, Water Breathing, Wind Wall}}
    
    \ability{Level 4:}{\emph{Confusion, Dimension Door, Dimensional Anchor, Evard's Black Tentacles, Fear, Fire Shield, Fire Trap, Hallucinatory Terrain, Ice Storm, Move Earth, Scrying, Shout, Solid Fog, Stone Shape, Wall of Fire, Wall of Ice}}
    
    \ability{Level 5:}{\emph{Cloudkill, Cone of Cold, Deltane's Fiery Tentacles, Fire and Brimstone, Fire Seeds, Greater Dispel, Dismissal, Fabricate, Lightning Leap, Mind Fog, Mirage Arcana, Nightmare Terrain, Prying Eyes, Seeming, Sending, Tactical Teleportation, Telepathic Bond, Transmute Mud to Rock, Transmute Rock to Mud, Wall of Force, Wall of Stone, Waves of Fatigue}}
    
    \ability{Level 6:}{\emph{Acid Fog, Chain Lightning, Contingency, Guards and Wards, Mass (Bull's Strength, Cat's Grace, Bear's Endurance), Mass Suggestion, Storm of Fire and Ice, Sunburst, True Seeing, Veil, Wall of Iron, Wall of Thorns}}
    
    \ability{Level 7:}{\emph{Banishment, Control Weather, Deadly Lahar, Earthquake, Forcecage, Greater Teleport, Incendiary Cloud, Mass Hold Person, Mass Invisibility, Mordenkainen's Sword, Phase Door, Planeshift, Prismatic Spray, Reverse Gravity, Waves of Exhaustion, Whirlwind}}
    
    \ability{Level 8:}{\emph{Deadly Sunstroke, Dimensional Lock, Greater Prying Eyes, Greater Shout, Horrid Wilting, Mass Charm Monster, Mass Polymorph, Prismatic Wall, Storm of Vengeance}}
    
    \ability{Level 9:}{\emph{Astral Projection, Elemental Swarm, Etherealness, Firestorm, Foresight, Mass Hold Monster, Meteor Swarm, Prismatic Deluge, Prismatic Sphere, Timestop, Towering Thunderhead, Wail of the Banshee}}
\normalsize

\spelllist{Altered Spells:}
\ability{Chain Lightning:}{this acts as a Lightning Bolt (see below), except that any time a target takes damage from it, another smaller bolt arcs out to a secondary target within 50 feet. The Warmage selects the target, who must not be within the area of the main Lightning Bolt. This requires a Ranged Touch Attack for each secondary bolt, and they deal half the damage of the primary bolt. If bouncing causes someone to be hit by the same bolt multiple times, they release multiple secondary bolts.}

\ability{Cone of Cold:}{This works like normal, except the entire area of effect is coated in ice (functioning as a Grease spell) and anyone who fails the save takes 1d6 Dexterity damage.}

\ability{Firestorm:}{This spell functions as normal, except the Duration changes to [Concentration], and the area of effect cannot be changed once chosen.}

\ability{Flame Arrow:}{This affects one target per caster level, for one round per level. All projectiles they throw or fire are treated as Flaming weapons.}

\ability{Lightning Bolt:}{When cast by a Warmage, Lightning Bolts may bounce off surfaces - either at a 90 degree angle either way, or 180 degrees straight back at the caster. The caster chooses each time. Each time the electric bolt bounces, the Save DC is reduced by 2, and the damage reduced by 5. If ever the Save DC or damage is reduced below 1 in this manner, it fizzles out. Targets must save separately each time they are hit by the same bolt, but the moment they pass a save, they automatically succeed on all future ones from the same bolt.}

\ability{Mass X:}{With the exception of Mass Polymorph, the spell functions as normal, but with one target per character level (before More Targets is applied), all within the appropriate range.}

\ability{Mass Polymorph:}{Two willing subjects per level are affected for up to one hour per level. They turn into Trolls, keeping their alignment, memories, personalities and allegiences, but replacing all other traits, even mental ability scores, with those of basic trolls.
This is a Transmutation [Polymorph] effect.}


\featname{Meteor Swarm}
\shortability{Evocation}{[Fire]}
\shortability{Level:}{Elementalist 9, Sorc/Wiz 9, Warmage 9}
\shortability{Components:}{V, S}
\shortability{Casting Time:}{1 standard action}
\shortability{Range:}{Long (400 ft. + 40 ft./level)}
\shortability{Area:}{Four 40-ft.-radius spreads; see text}
\shortability{Duration:}{one minute or until discharged}
\shortability{Saving Throw:}{None or Reflex half; see text}
\shortability{Spell Resistance:}{See below}
\noindent One meteor per four caster levels, each a foot in diameter and surrounded by flame, whirls around the caster. With an Attack Action, the caster may unleash one meteor out to Long Range, requiring a Ranged Touch Attack. If it hits, the target takes 10d6 Bludgeoning damage with no Spell Resistance or Saving Throw. Whether or not it hits, the meteor detonates into a 20' radius burst, dealing 1d6 Fire damage per caster level with a Reflex Save for half. Spell Resistance applies against this, however if the meteor hits the target, they automatically fail the saving throw and do not benefit from Spell Resistance. The entire area is then reduced to Difficult Terrain.

\smallskip\noindent Furthermore, while the spell is ongoing, when the caster is hit by a melee or ranged attack that requires an attack roll, he may sacrifice one meteor without using an action of any kind, to reduce the incoming effect. Roll the damage (both the 10d6 Bludgeoning and the fire damage), and subtract the total from the damage of the incoming attack, taking the remainder as normal. If the damage is reduced to or below zero, the entire effect, including any special effects, is negated. Note that attacks that deal zero damage are automatically negated in this way.