\subsection{Formians: In Many, One}
%By Quantumboost

\quot{``There is no good for the bee that is bad for the hive.''}

\desc{The Formians are the current Lawful Outsiders, and true to their name are giant ants -- and that's most of what you need to know about them. Formians represent Law as Conformity -- conformity to their society, and removal of anyone who won't conform. They're expansionistic, militaristic, and each hive effectively acts as one organism as they spread across the gears. Like most portrayals of giant ant-people, Formians are telepathic and have a hive-mind. However, they are giant ant \emph{people}, and each individual Formian above Worker level has a level of mental ability at least what you might expect from a human. So interacting with them is a lot like working with the Borg - you can totally }

\desc{The most straightforward way to use the Formians is to have them show up and want to conquer your favorite hangout -- either taking away the people to work in their hive (which is a lot like what the Hobgoblins will do, except that there probably isn't any chance of advancement from the slave thing) or turning it into a new hive, superintelligent mind-enslaving queen ant and all. But you can also use them as more mundane antagonists or even as potential allies. Unlike Modrons, Formians actually have discernable goals and motivations, their motivations just all happen to be ``what is good for the hive''. If you're working towards a goal or offering a reward that they think will help the hive, they'll totally be willing to work with you and might not even bother you about the whole assimilation thing beyond handing out some pamphlets.}