\classname{Mechanus Warrior} \label{class:mechanuswarrior}
\vspace*{-8pt}


\begin{table}[tbh]
\begin{small}
\begin{tabular}{lp{3cm}p{0.7cm}p{0.7cm}p{0.7cm}l}
Level  &Base Attack Bonus &Fort Save &Ref Save &Will Save &Special\\
1st &+1 &+2 &+0 &+0 &Weapon Cogs, Mechanus Mind, Lawful Aura, Iron Guts, Track, Construction\\
2nd &+2 &+3 &+0 &+0 &Endurance, \textit{locate object} \\
3rd &+3 &+3 &+1 &+1 &Strength Enhancement, Fortification \\
4th &+4 &+4 &+1 &+1 &Tool Cogs, Vision Implants \\
5th &+5 &+4 &+1 &+1 &Wounding, Constitution Enhancement, \textit{clairvoyance}\\
6th &+6/+1 &+5 &+2 &+2 &Stability\\
7th &+7/+2 &+5 &+2 &+2 &Keen, \textit{locate creature}, Fortification\\
8th &+8/+3 &+6 &+2 &+2 &Assistant Cogs\\
9th &+9/+4 &+6 &+3 &+3 &Knockdown, Planar Tracking, Intelligence Enhancement\\
10th &+10/+5 &+7 &+3 &+3 &Iron Body, \textit{legend lore}\\
11th &+11/+6/+6 &+7 &+3 &+3 &Vision Implants, Fortification\\
12th &+12/+7/+7 &+8 &+4 &+4 &Greater Weapon Cogs\\
13th &+13/+8/+8 &+8 &+4 &+4 &\textit{discern location}\\
14th &+14/+9/+9 &+9 &+4 &+4 &Instability\\
15th &+15/+10/+10 &+9 &+5 &+5 &Mechanus Cannon\\
16th &+16/+11/+11/+11 &+10 &+5 &+5 &\\
17th &+17/+12/+12/+12 &+10 &+5 &+5 &\\
18th &+18/+13/+13/+13 &+11 &+6 &+6 &\\
19th &+19/+14/+14/+14 &+11 &+6 &+6 &\\
20th &+20/+15/+15/+15 &+12 &+6 &+6 &\\
\end{tabular}
\end{small}
\end{table}

Skills: 4 + Int Bonus per level: Appraise, Climb, Concentration, Craft, Decipher script*, Disable Device, Forgery*, Knowledge (Any), Move Silently, Search, Sleight of Hand, Speak Language, Spellcraft, Spot, Survival
    *: a Mechanus Warrior automatically has this skill at max ranks.



Armor and Weapon Proficiencies: The Mechanus Warrior is proficient with light, medium, and heavy armor, a well as with shields, tower shields, Mechanus Armor, Golem Armor, and Omega Armor. The Mechanus Warrior is proficient with Simple and Martial weapons as well as all forms of Crossbow, the Spiked Chain, the Buckler Axe, the Maul, the Dire Pick, the Dire Flail, the Heavy Poleaxe, the Mancatcher, and his own Weapon Cogs.


\ability{}{}
Weapon Cogs (Ex): The Mechanus Warrior has a special relationship with minor Modron assistants that live inside his body called Cogs. These creatures transform into useful items. At first level, the Mechanus Warrior can transform them into various self propelled weapons that are attached to his body or withdraw them into himself as a Move Action. A Mechanus Warrior can always use their Strength Modifieron attack rolls with their Weapon Cogs, regardless of whether it is a ranged weapon or melee weapon. Attacking with a Weapon Cog only literally requires one hand to use it, but the other hand cannot make a weapon attack in a Two Weapon Fighting sort of way. Weapon Cog weaponry always has an enhancement bonus of 1/3 the character's level to attack and damage, rounded up (no max). Available weapons are as follows:

    Gear Thrower. This is like a crossbow that never needs to be reloaded and has a very high rate of fire. It has a range increment of 40' + 5' per level, and every time it is fired the Mechanus Warrior can make two attacks on the same target (or on two different targets if they are within 10' of each other). Each attack does 1d6 of slashing damage with a threat range of 18-20.

    Rotary Cutter. This is like a buzz saw. It is a melee weapon that does 2d6 of slashing damage with a critical multiplier of x4.

    Magnetic Dust. This weapon is a series of extremely fine gears held together with magnetic force. It appears vaguely sword or ax shaped, but is gray and semi-transparent. It does 1d8 of piercing damage with a x2 critical, but attacks made with it are melee touch attacks.

    Diamond Drill. This weapon is a drill. It does 2d6 of piercing damage with a x4 critical. While it is a melee weapon, Strength bonus is not added to damage. But the Drill bypasses any DR or Hardness.



Mechanus Mind (Ex): A Mechanus Warrior may make Fort saves instead of Willpower saves against Illusions. A Mechanus Warrior may use their Intelligence Modifier instead of their Wisdom Modifier on skill checks.

Lawful Aura (Su): A Mechanus Warrior responds to detect law and similar effects as a Lawful Outsider or Cleric of their Hit Dice.

Iron Guts (Ex): A Mechanus Warrior is full of metal bits, and is immune to any poison that does not affect constructs.

Track: A Mechanus Warrior gains Track as a bonus feat. Also they can find and disable traps with no caps to the difficulty they can attempt.

Construction: A Mechanus Warrior may use their highest ranked Craft Skill in lieu of any other Craft skill for any purpose. In effect, the player need only write "Craft" rather than "Craft Blacksmithing" (or whatever). In addition, the Mechanus Warrior may use their ranks in Craft as a Caster Level for purposes of qualifying for and using Item Creation feats.

Endurance: At 2nd level, a Mechanus Warrior gains Endurance as a bonus feat.

Spell-like abilities (Sp): As a Mechanus Warrior gains levels, they get access to a number of divinations as spell-like abilities. Each spell can be invoked once per day for every attack guaranteed them by their BAB (so they get a second use of each per day at BAB +6 and a third at +11). At 2nd level they can use locate object, at 5th level clairvoyance, at 7th level locate creature, at 10th level legend lore, and at 13th level discern location.

Strength Enhancement: Starting at 3rd level, the Mechanus Warrior has an Enhancement Bonus to Strength. This bonus is 1/3 of their level, rounded up (no max).

Fortification: As a Mechanus Warrior becomes more powerful, their insides become more redundant and mechanical. At 3rd level, this manifests as a 25% chance to negate a critical or precision based damage boost against them - as light fortification armor. At 7th level this increases to a 50% chance to negate, and it rises to 100% at 11th level.

Tool Cogs (Ex): At 4th level, the Cogs can become all manner of useful devices, and with a Move Action can be converted into masterwork tools appropriate to any task. They retain their enhancement bonus of 1/3 level, rounded up, for this purpose.

Vision Enhancement (Ex): At 4th level, a Mechanus Warrior's cybernetic eyes are of such power that he can see in darkness and cannot be dazzled or blinded by flashing light. Not even magical flashing light. At 11th level they improve further, and vision is no longer impaired by fog, smoke, or flying sand. And the Mechanus warrior sees the true form of things within 120' as if under the effects of true seeing.

Wounding (Su): From 5th level on, when a Mechanus Warrior damages a creature with a Weapon Cog attack, the target also suffers 2 points of Constitution damage.

Constitution Enhancement: Starting at 5th level, the Mechanus Warrior has an Enhancement Bonus to Constitution. This bonus is 1/3 of their level, rounded up (no max).

Stability (Ex): At 6th level, a Mechanus Warrior can stand and move in difficult terrain without penalty and trip or grapple attempts against him have a -4 penalty.

Keen (Su): At 7th level, a Mechanus Warrior's Weapon Cogs attacks have their threat range doubled.

Assistant Cogs (Ex): From 8th level on, a Mechanus Warrior is constantly assisted by cogs in various shapes. Whenever a Mechanus Warrior makes a d20 roll, they get a +2 Aid Another bonus.

Knockdown (Su): When a 9th level Mechanus Warrior damages a target with a Weapon Cog attack, the target must make a save with a DC of 10 + 1/2 Level + Strength Modifier or fall prone. Bonuses against Trip attempts add to this saving throw.

Planar Tracking (Su): When a 9th level Mechanus Warrior is tracking a creature and comes to a point at which they have teleported or traveled through dimensions, they can follow them with a DC 25 Survival check. Any required portal or rift is briefly reopened, and they can take up to one extra medium creature per level with them.

Intelligence Enhancement: Starting at 9th level, the Mechanus Warrior has an Enhancement Bonus to Intelligence. This bonus is 1/3 of their level, rounded up (no max).

Iron Body: At 10th level, mechanus steel appears even on the skin of the Mechanus Warrior. They gain a +4 bonus to their Racial Bonus to Strength, and their native form has a +6 Natural Armor bonus.

Greater Weapon Cogs (Ex): From 12th level on, the Mechanus Warrior can make their cogs into more extreme and personalized weapons. Go ahead and make some custom ones that are kind of like this:

    Chain Gun. It's a series of spinning tubes, out of which a vast number of steel needles fly out. Each attack attacks all creatures in a line out to Long Range, and while it does not inflict criticals at all, it strikes each target once for each point the attack beats the AC by. Each hit in simply 1d6 of Piercing Damage, but remember that it's a Weapon Cog.

    Steam Hammer. It's like a giant metal hammer that is accelerated by a jet of steam whenever it is swung. It is a reach weapon that inflicts 3d8 of Bludgeoning Damage with a x4 critical. Every time it hits, there is a burst of steam so severe that it impairs visibility like obscuring mist, dissipating in 2 rounds.

    Thresher Combine. It's a bunch of scythe blades that spin around the Mechanus Warrior, grab things, and tear them to pieces. When the user attacks, they may make an attack on every opponent within reach. Each hit does 2d6 of Slashing Damage with a x4 critical, and any target damaged by it is also entangled and considered to be in Difficult Terrain until they begin a turn outside the Cog's threatened area.

    Furnace Blaster. It's like a large brass lantern, and when it is opened a jet of fire lurches out to reduce things to burning slag. As an attack action, the Mechanus Warrior can spray 10d6 of Fire Damage in a cone out to short range. Creatures in the area may attempt a Reflex save for half damage, DC 10 + 1/2 Level + Strength Modifier.



Instability (Su): At 14th level, the Mechanus Warrior can shift between the material and the ethereal at will. They can go to the Ethereal plane whenever they want, but they can also phase in and out like blink, but without themselves suffering any miss chances.

Mechanus Cannon (Su): A 15th level Mechanus Warrior can open a gate to places they have been in Mechanus. They can use this for travel to that plane, but they can also open one up right in front of one of the great pistons of that plane just long enough for a titanic column of metal to come down and crush things. The Mechanus Warrior nominates a 20' diameter circle and everything in that circle up to 100' off the ground is pounded by a giant column of Mechanus Steel. This does 20d6 of nonmagical Bludgeoning Damage and there is no save.
