\chapter{High Adventure on the Cordant Planes}
\quot{``Then you shall face your punishment in a realm of neither the Heavens nor the Hells...''}

The Cordant Planes are extremely different in character from the Upper and Lower Planes, just as the difference between Law and Chaos is fundamentally different from that between Good and Evil. Out of the Tomes, the Book of Gears deals most heavily with one of the Cordant Planes -- Mechanus -- so really this is the best place to handle them.

\section{High Adventure in... Mechanus!}

Mechanus is the plane of Ultimate Law, and while the actual nature of Law is dependent on how you handle Alignment in general, Mechanus is always giant clockwork land filled with robots and ant-people. Currently the official ``robots'' are the Inevitables, but you might very well have Modrons. The below examples assume

%\input{adventuremechanus}

\section{High Adventure in... The Outlands!}

The souls who end up in The Outlands are those who are ``Neutral'' or follow one of the dieties who live here. Given how alignments in 3.0 work, this means basically everyone who doesn't choose a particular other alignment and it \emph{also} includes druids and other ``balance nuts.'' The Outlands is a lot like the Prime Material Plane in that there is a bunch of nature and also a bunch of cities, except all of it is made of soul-stuff and magic cuts off near the center. Generally you'll be here either because you're traveling between the other planes via the gate-towns on its rim or because you have business with the Mind Flayers, Naga, Obad-hai, or any other of the random domains which exist around here.

%\input{adventureoutlands}

\section{High Adventure in... Limbo!}

Limbo is a land of Chaos, and like with the rest of the Law-Chaos fiasco it's not entirely clear what this means. Mechanically speaking, Limbo is made of a bunch of ``Chaos Matter'' which is for some reason shaped entirely by the minds around it and otherwise randomly cycles between the basic elements. Also the natives are all supernatural Giant Frogs or Githzerai. Despite the unstable nature of everything in Limbo, there are a few areas of interest which are either inhabited (and thus kept in a stable form pretty much all the time) or some sort of psychic storm.

%\input{adventurelimbo}