\chapter*{Preface to the Book of Gears}

\desc{Cooperative Storytelling is essentially all about artifice. The stories we create are created, the shared narrative is an illusion which fills our mind and pushes us forward. So it is no surprise that creating things within that narrative is so very contentious. Building a house in the game or creating an illusion in the story is a}

\desc{An illusion is something that isn't real inside a story that isn't real. Forging a sword is creating something within a tale that is being created around it. These actions, while very integral to the source material upon which our cooperative storytelling games are based, are yet one more step removed from reality when contrasted with the old standards of pretending to be a knight who kills imaginary dragons to save fictitious princesses.}

\desc{So it seems not at all surprising in retrospect that the rules we have used to represent the creation of stuff within the game world have historically been extremely unsatisfactory. Creating things takes time, which is a problematic concern in a game where time passages narratively. That means that the time a character spends nailing boards together for his dream house may be spent in a montage that ends in subtitles reading ``six months later'' and it may happen interspersed with a rollicking adventure where seconds count and the hammering essentially never gets done.}

\desc{The result has been that previous editions have attempted to put additional or alternate costs on crafting of all sorts. From Constitution points to years off your life to XP, D\&D has experimented with about a dozen different rubrics by which characters could trade one part of their character for more magic items. In almost all cases this allowed players to trade things they weren't using anyway for powerful artifacts that allowed them to conquer worlds, although in a few cases the flip side showed up and made item creation so crappy that people seriously didn't do it at all. Needless to say, this has been unsatisfying, and it is our intention to help remedy these problems.}

\desc{The rules presented here present a different take entirely. Creating magic items is something that takes only time, and adventures can be expected to be completed without ever doing it at all...}