\section{Dragons: The Hoarding}

\noindent\desc{From the aggregate collection of human mythology, dragons are monsters with scales. They've been both the size of a dog to the size of a mountain, their leg number is as varied as the tastes of a hedonist, their powers as different as what humans are given. About the only seeming constant is that they're scaly and their power is part of their natural state, as opposed to the tool-using nature of humans.}

\noindent\desc{Now, what and how dragons are portrayed in the past only matters if you're trying to recreate mythology. Modern dragons have a distinct Western root. It's been changed since then, just look at historical depictions of Saint George's dragon, that thing is the size of a mastiff. The same source material depicts them as clever beasts at best, and most just had super-poison (fitting the serpent theme). As time passed, they got smarter, bigger, firey, wealthier, and more magic-y. It's gotten to the point where we just don't accept dragons smaller than draft horses as terribly threatening. Ultimately, the closer you resemble Tolkien's Smaug, the more draconic you are.}

\noindent\desc{In the RPG demographic, they're subject to fanboy wank. Even the 'mook' dragons you're allowed to kill are artificially made tougher than their CR would indicate (the infamous [Awesome] subtype), then there's the not uncommon Draconis Ex Machina. We largely don't like them smaller than a horse unless they're basically hatchlings, and yet we want them as final bosses from the dungeon crawl from the very beginning so we can live up to the title of the game, which doesn't happen under the current rules (medium+ CR 4 dragons, impossible!).}

\noindent\desc{We've ideally seperated gross material wealth from power in the Book of Gears, but this still leaves the problem that barring fanfic dragons, they largely don't use magic items and run around in the buff. We need to have six or seven of their magic item 'slots' be taken up with their own body (likely all minor). This way, their interest in baubles, while present, will largely sit in their Bat Cave because they can only use one at a time. It also sets us up with the precedent for their corpse retaining mojo, be it using their scales for magical dragon armor, their third eye for crystal balls, their teeth for swords/daggers, their blood for potions, etc. Even if you don't use that method, their pile of gold can still hold an array of magic items that were understandably unused by the dragon.}

\noindent{This still leaves the big issue, their hoard. Why do they amass such material wealth they're unwilling/unable to spend while also encouraging treasure-hunting adventurers to kill them? There have been several excuses made for it. I've heard of them using it as a sanctuary-style focus to help them mature and age into their magical power at a rate not dissimilar to wizards, which is a viable method of turning wealth into power. I've also heard the idea them being like magpies, and their intelligence makes this instinct more refined.}

\input{baseclasses/halfdragon}