\quot{``And that is why I had to make things so difficult for you, daughter. I simply had to see if you'd live through it.''}

Many of the fey are roughly human sized, able (though not necessarily willing) to wear human clothes, able to use human weapons and tools, and all in all pretty capable of performing human actions in response to standard adventuring challenges. It is notable then, that despite the substantial \emph{demand} for fairy player characters and the basic playability of the \emph{concept}, that there isn't anything in D\&D that is actually game mechanically playable. Dryads, Feytouched, Pixies, Nixies, Satyrs, and Sirines are all in high demand as PCs, and only overly harsh rules on playing ``nonstandard'' PCs stand between these players and the characters they want to play. What follows is a reimagining of some of the most playable Fey into rules that are game mechanically as well as conceptually playable within a D\&D game.

\subsection{The Fey Touched}
\quot{``All my life, I have never fit in. Not in town, not in the forest. In some integral fashion I am unlike those around me, and I believe it is my fate to live and die alone.''}

\noindent\desc{WotC has made no secret of the fact that they like doing bad things to the bad touched races. But if there was one Bad Touched race that is \emph{almost} worth the level allowance, it would be the Fey Touched. Every one of them looks completely different, but they all have the same power set: spell-like \spell{charm}, and of course an Immunity to Mind Affecting effects. That goes a long way to being worth something, though of course there is no way you can survive as a character with a Level Allowance and a Constitution penalty -- the very idea is absurd. So of course, the race has to be overhauled, because that just isn't reasonable.}

\noindent\desc{Now if you're one of the people who wonders why a product of fairies and humans, who both conspicuously lack an immunity to mind affecting magic, would have an immunity to mind affecting magic -- you aren't alone. That question comes up about as often as any other with regards to the fey touched. Of course, not all of those born to fey and human stock are immune to mind affecting magic, as you might expect from a group so diverse that some have bug parts and others are simply beautiful humans, while still others look like crazy rock men with teeth sticking out all kinds of places, the powers that a fey-touched is born with are extremely random. The powers of the fairies are more than a little bit chaotic in nature, and no two babes born to these couplings are the same. Unfortunately, these mulish offspring are also \emph{interesting} both in the general sense and, much more to their detriment, to other fairies in particular. The unmitigated interest of the fey is hard on a small child, so fey touched who are not immune to compulsion effects are going to find themselves at the bottom of a pond or jumping out of a tall tree long before they reach adulthood. Indeed, feytouched immune to compulsion effects are the only ones that ever reach maturity -- the well meaning but deadly interest of the fairy family members simply weeds out any other possible results.}

\noindent\desc{That's not an excuse for the package presented in the Fiend Folio as a whole, that's simply unplayable. But it's close. Here's our version:}

\textbf{Feytouched}

\listone
\item Fey Type
\item 30 foot movement rate
\item Low-Light Vision
\item +2 Dexterity, +2 Charisma, -2 Constitution. Feytouched are graceful and those which are not beautiful are terrifying, but they are fragile like flowers.
\item Immunity to [Compulsion] Effects
\item Magic Affinity: Every Feytouched is different, and marked by the signature magics of the fey in a different manner. Every Feytouched has one spell that can be used once per day as a spell-like ability. This spell is chosen at 1st level and cannot be changed. Any 1st level Illusion or Enchantment spell from the Sorcerer/Wizard list is fair game, and the sve DC is Charisma-based.
\item Favored Class: Bard
\item Feytouched speak Common and Sylvan. Bonus Languages may be selected from the following list:\\
Aquan, Auran, Elvish, Draconic, Dwarvish, Druidic, Goblin, Gnoll, Gnome, Halfling.
\item Level Adjustment: +0
\end{list}