\begin{multicols}{2}

\hypertarget{feat:weaponrighteous}{}\babfeat{Weapon of Righteous Destruction [Combat]}{

Your hands make whatever is being held by them holy and on fire. For some reason this doesn't make them melt or burn up.}{

Whatever weapon you are wielding is considered Magical (+1/3 bonus/level) in addition to any other properties that it has. Your unarmed attacks, even if not proficient, count for this effect.}{

The above, Flaming weapon.}{

The above, Holy instead of Flaming.}{

The above, Sun weapon, Fort save. (BoG)}{

The above, Vorpal weapon (BoG).}

The following feats from the Tome of Fiends are now also [Celestial] feats, and you don't pick up fiendish traits for taking them as a [Celestial] feat:

Breath Weapon 

Elemental Aura 

Extra Arms 

Extra Summons 

Greater Teleport 

Harmless Form 

Large Size 

Huge Size 

Heighten Spell Like Ability 

\descfeat{Product of Celestial Dalliance}{One of your recent ancestors was an celestial Outsider or from a good-aligned plane. Maybe your parents play it off as a virgin birth, maybe your dad became a Saint.}
You may take any [Celestial] feat, and gain Resistance 5 to Acid, Cold, Electricity. You also gain one of the Angel, Archon, Eladrin, or Guardinal subtypes, and a Smite Evil attack usable at will that does bonus damage equal to 1/2 of your strength modifier.

\featname{Wings of Good [Celestial]}
You gain wings and a fly speed equal to double your base land speed with good maneuverability that requires special armor to stay aloft. These must be feathery or energy-based.

\hypertarget{feat:lordofdeath}{}\skillfeat{Lord of Death [Necromantic] [Skill] [Leadership]}{
A whole bunch of skeletons and crap show up to fight under your tattered banner.}{
Knowledge (Religion) Ranks:}{
You have a Command Rating equal to your ranks in Knowledge Religion divided by five (round up). You are a Necromantic leader (see Heroes of Battle).}{
You can muster a group of followers. Your leadership score is your ranks in Knowledge Religion plus your Wisdom modifier. Your followers are all mindless Undead. You don't make them or anything, they just show up.}{
You are able to delegate command to a loyal cohort. Your cohort is an intelligent and loyal Undead creature with a CR at least 2 less than your character level. Cohorts gain levels when you do.}{
Your followers swell in number to that of an army.}{
Your allies gain energy resistance to Positive Energy equal to your level while they are within line of sight of you.}

\hypertarget{feat:tyrant}{}\skillfeat{Tyrant [Skill] [Leadership]}{
You push people around and get larger and larger groups trapped in the iron gauntlet of your brutal rule.}{
Intimidate Ranks:}{
You inspire such terror that creatures you intimidate continue to act intimidated after you leave, too afraid to raise their voice in defiance even after you have apparently left them far behind.}{
You can muster a group of followers. Your leadership score is your ranks in Intimidate plus your Strength modifier.}{
Your followers swell in number to that of an army.}{
Your mere presence inspires fear and can break a battle. Enemies with more than 5 hit dice less than you do must make a Will save (DC 10 + ½ Level + Strength Modifier) of flee in \emph{panic}.  This is a [Fear] effect.}{
Your presence causes despair in even brave opponents. All enemies within 30' of your suffer a -2 Morale penalty to Willpower saves.}

\hypertarget{feat:monsterrancher}{}\skillfeat{Monster Rancher [Skill] [Leadership]}{
You can breed and train a large number of crazy beasts.}{
Handle Animal Ranks:}{
You can use Handle Animal as if it were Diplomacy when dealing with Magical Beasts and Dragons. You can do similarly with Aberrations and Plants with an Intelligence Score that is less than 9.}{
You can muster a group of followers. Your leadership score is your ranks in Handle Animal plus any synergy bonuses you gt to that skill. Your followers can, and must be monsters.}{
You have a loyal cohort that is a monster of some kind. A cohort is an intelligent and loyal creature with a CR at least 2 less than your character level. Cohorts gain levels when you do.}{
You know what any monster is unless it is disguised by illusion, and you can look up its stat line in the appropriate monster book when devising your strategies.}{
Once per day, you can reroll a saving throw allowed by a Supernatural Ability.}

\hypertarget{feat:armyofdemons}{}\skillfeat{Army of Demons [Skill] [Fiend] [Celestial] [Leadership]}{
You have an army of planar crazy crap.}{
Knowledge (Planes) Ranks:}{
You have a Command Rating equal to your Knowledge: Planes ranks divided by five (round up).}{
You can muster a group of followers. Your leadership score is your ranks in Knowledge: Planes plus your Charisma mod. These followers can and must be outsiders.}{
Your followers swell in number to that of an army.}{
You own a planar stronghold.}{
Your allies gain a +2 morale bonus to all saving throws if they can see you and you are within medium range.}

\hypertarget{feat:bureaucrat}{}\skillfeat{Bureaucrat [Skill] [Leadership]}{
You have a functioning guild that makes stuff for you and gives you money.}{
Appraise Ranks:}{
You draw an income for working as an administrator, getting 1 GP/week per rank in Appraise.}{
You can muster a group of followers. Your leadership score is your ranks in Appraise plus your Intelligence modifier. These followers all have profession and craft skills.}{
You get your own Stronghold.}{
You get a +2 bonus to profit checks.}{
Your guild goes planar, your number of followers swell to the size of an army and their ranks start filling up with producers and managers from other planes of existence.}

\end{multicols}