\subsection{Modrons: Singularity of Purpose}
% By Frank Trollman

\desc{For those of you who don't remember: Modrons are the original creatures of Law from the old days of AD\&D. They haven't been seen very often because they were originally written as a joke. Their very existence is as offensive to many players as the fact that they were essentially retconned out of existence is to others. And what's that all about? It's because the Modrons were originally written up as giant dice. Yes, really. The different types of basic Modron are shaped like four sided dice, six sided dice, 8 sided dice, the whole thing.}

\desc{So if your DM jumps on the ``let's forget this ever happened'' bandwagon, we understand. The original write up of the Modrons was actually pretty insulting. But since then there have been a number of variously successful attempts to rehabilitate them and make them independently awesome. Different Modron art has been made by Tony DiTerlizzi and Eric Campanella that looks pretty darned awesome – and not like your DM put a 6 sided die on the battle mat at all. Instead each Modron looks like a ghastly hybrid of metal and flesh covered with cogs and wheels where spindly appendages emerge from a solid (though not rollable) core.}

\desc{So assuming that you use some of the reform Modrons from late in 2nd Edition, the Modrons are actually pretty cool. They represent the idea of Law as an implacable and incomprehensible force. They are at their best when portrayed as being so single mindedly focused on some long term goal that they actually don't even care about you. Sometimes they destroy your village, sometimes they don't, and there's really no predicting that sort of thing unless you're knowledgeable about the Big Plan. Now I know what you're thinking… that having a plan so convoluted and far ranging that mortal minds cannot grasp it or predict its unfolding is actually indistinguishable from not having a plan at all and just performing actions at random. And yeah… that's true. That's DnD alignment for you.}

\desc{The Modrons come from a city in the Clockwork Nirvana called Regulus and have a rigid caste system where more powerful Modrons are told more of ``the plan'' than less powerful Modrons are told less. Each Modron is told exactly as much as it needs to know to complete its assigned tasks. And in the face of a long term plan of this magnitude, that pretty much means that every Modron is kept entirely in the dark about just what the heck it is doing or why it is doing it. The Modrons are arranged into a rigid caste system with no possibility (or concept) of personal advancement. That would be pretty stultifying if they were like humans where they all started out equal, but they aren't like that at all and Modrons of each caste sincerely don't have any desire to move up to another caste. At the very bottom (or at least, ``most numerous and least clued in to the plan'') there are the cogs. Cogs look like little gears and have the ability to transform into any object of roughly a cubic foot or less. When properly supervised by a higher level drone, they can take the shape of quite complicated pieces of clockwork and frequently do so. Above them are various drones of various shapes and sizes, each constructed of materials mechanical and biological to fulfill its role in the great plan.}

\desc{At the top there is Primus, who has been variously described as everything from an Intelligent Item to a reasonably powerful Outsider to a guy working the machinery behind a curtain. Seriously, your campaign can potentially reveal anything you want it to about what is really in the center of Regulus because that's been retconned so many times that noone knows what the official answer even is right now. Primus has been killed off several times in official continuity as part of various authors attempting to delete the Modron race from D\&D. However, someone always brings Primus back, because D\&D never really throws anything away (except the pygmies, that was too racist even for the 1980s). If you demand continuity in your life, then it seems that Primus simply can be killed time and time again, each time getting replaced with a new Primus who is in turn enlightened as to the nature of the big plan and granted the authority to turn the wheels of Regulus.}

\desc{The Modrons mostly sit around and operate the machinery in Regulus that apparently keeps all the giant gears and pistons running in the Clockwork Nirvana of Mechanus. So even if they exist in your game's continuity, chances are good that you'll never encounter them. Every so often, a whole lot of Modrons open up gates to other planes and start wandering around in big groups doing… stuff. The amount of time that Modrons spend between their ``March'' is supposed to be constant but actually every single Modron March that has ever been mentioned in any published adventure or story has taken place out of sequence so one is forced to conclude that actually the Modron March happens whenever the big plan calls for it and rumors of a great schedule are just rumors. The Modron March is generally not harmful and the Modrons don't seem to go out of their way to chase anyone who gets out of the way, so it doesn't seem to be an invasion. Although who knows? The plan of the Modron is so incredibly far reaching that it is entirely possible that they simply walk around in huge armed groups wandering seemingly aimlessly through the planes at irregular intervals doing no harm to anyone so that at some later date they can do the same thing and then just destroy some enemy that is predicted in the distant future.}

\desc{Unfortunately, now I'm going to have to talk about the Hierarchs. These are a layer of bad asses that live in the eco niche between drone and god. They are roughly equivalent to the high end fiends and celestials – coming in various flavors and power levels like Mariliths and Balors do for the demons. Unfortunately, noone has ever overhauled the art on these guys to the point where they don't look like ass. Sorry, the Hierarchs of Regulus look like they were drawn by Napoleon Dynamite and there's nothing nice I can say about them. If I were personally inclined to run with these guys at all I would be forced at gun point by the players to have them look like… something else. Putting them back into the theme of the Modron Drones is probably best, because at least then they appear to be Modrons in the same way that a Trumpet Archon is readily identifiable as an Archon. Which means that honestly the Hierarchs are much cooler if they look like Daleks. And the truth of that statement is probably the single greatest argument for walking away from the whole thing, as advised by Monte Cook.}

\desc{So what do Modrons do in a story? Mostly they show up with a specific set of instructions that they attempt to fulfill. They speak their own language but the more powerful ones also speak additional planar languages (hope you speak Formian). If the player characters attempt to prevent the Modrons from doing their thing, the Modrons will fight. Otherwise they'll simply complete their task and leave. If other creatures attempt to stop the Modrons, they'll fight them. Modrons hold no grudges and have no loyalty but to the plan. They will seamlessly switch sides in a battle if a different group proves more detrimental to their mission. The Modrons don't know why they are doing what they are doing, only Primus does (assuming Primus exists, in some versions The Plan is actually a flaw in Mechanus and there is no reason for any of it). And yet they will fight to the death to complete their mission. With good art and weird dialogue, the Modrons can serve as dynamic antagonists or useful allies.}

\desc{Like an amazingly intricate puzzle box, the Modron Race unfolds, collapses, and progresses in countless ways, but at the end, it's still a puzzle box. The entire point of the Modrons is that you never get clued in to what they are doing or why they are doing it. But if you don't want to use them at all, I understand.}